---
title: "Motif Heatmap"
output: html_notebook
---

This tutorial can be used to produce the motif heatmaps. In this tutorial, the current Seurat object has many clusters, all of which are labeled with their cell identity. The objective of this tutorial is to produce two heatmaps, one of the top 15 motifs of each EC cluster and one with the top 20 motifs.


### Set Values for Project

```{r} 
object.name = "object.rds"
cell.type = "EC"
```  

### Necessary Packages

```{r} 
library(Seurat)
library(Signac)
library(viridis)
library(ggplot2)
```

### Useful Functions

```{r} 
# Funtion to simplify isolating the EC clusters
GetGroups <-function(
  object,
  group.by,
  idents
) {
  if (is.null(x = group.by)) {
    obj.groups <- Idents(object = object)
  } else {
    obj.md <- object[[group.by]]
    obj.groups <- obj.md[, 1]
    names(obj.groups) <- rownames(x = obj.md)
  }
  if (!is.null(idents)) {
    obj.groups <- obj.groups[obj.groups %in% idents]
  }
  return(obj.groups)
}


# Function for producing heatmap
ProduceHeatmap <-function(
  object,
  number.motifs = 15,
  cell.id,
  fold.enrichment = 2.8
) {
  # Objects which will be used later
  matrix_list <- 1
  top_genes = c()
  
  # For each EC cluster...
  for(idx in c(1:length(cell.id))) {
    message(paste0('- Getting fold values for ', cell.id[idx]))

    # Find the motif peaks relative to the other EC clusters
    da_peaks <- FindMarkers(
      object = object,
      ident.1 = cell.id[idx],
      only.pos = TRUE,
      test.use = 'LR',
      latent.vars = 'nCount_peaks'
    )
    
    # Test the differentially accessible peaks for overrepresented motifs
    enriched.motifs <- FindMotifs(
      object = object,
      features = rownames(da_peaks)
    )
    
    # Sort by P-Value and then motif FC
    enriched.motifs <- enriched.motifs[ -enriched.motifs[, 6], ]
    
    # Edit dataframe
    row.names(enriched.motifs) <- enriched.motifs$motif.name
    
    # Keep top n motifs
    top_genes <- c(top_genes, row.names(head(enriched.motifs, n = number.motifs)))
    enriched.motifs <- enriched.motifs[, c(6, 8)]
    names(enriched.motifs)[1] <- cell.id[idx]
    
    # Merge all dataframes together
    if(matrix_list == 1){
      matrix_list <- enriched.motifs
    } else {
      matrix_list <- merge(matrix_list, enriched.motifs, all = TRUE)
      matrix_list[is.na(matrix_list)] <- 0
    }
  }
  
  # Remove repeating motifs
  top_genes_unique <- unique(top_genes)
  
  # Only keep significant motifs
  matrix_list_unique <- matrix_list[matrix_list$motif.name %in% top_genes_unique,]
  
  # Make motif names the row names to make the matrix numeric
  row.names(matrix_list_unique) <- matrix_list_unique$motif.name
  matrix_list_unique <- matrix_list_unique[, -c(1)]

  # Prepare matrix to make heatmap 
  matrix_list_unique$current <- paste0('EC', 1)
  new_mat <- data.frame(
    rownames(matrix_list_unique), 
    matrix_list_unique$current, 
    matrix_list_unique$EC1
  )
  matrix_list_unique <- matrix_list_unique[, -c(1)]
  colnames(new_mat) <- c('X','Y','Z')
  for (i in c(2:8)) {
    matrix_list_unique$current <- paste0('EC', i)
    temp <- data.frame(rownames(matrix_list_unique), matrix_list_unique$current, matrix_list_unique[,1])
    colnames(temp) <- c('X','Y','Z')
    new_mat <- rbind(new_mat, temp)
    matrix_list_unique <- matrix_list_unique[, -c(1)]
  }
  colnames(new_mat) <- c('X','Y','FoldEnrichment')
  new_mat$X <- factor(new_mat$X,levels = rev(top_genes_unique))

  # Make heatmap
  new_plot <- ggplot(new_mat, aes(Y, X, fill=FoldEnrichment)) + 
    geom_tile() +
    scale_fill_viridis(rescaler = function(x, to = c(0, 1), from = NULL) {
      ifelse(x<fold.enrichment, scales::rescale(
        x, to = to, from = c(min(x, na.rm = TRUE), fold.enrichment)
      ), 1)
    }) +
    theme_bw() +
    ggtitle(print0("Top ", number.motifs, "Motif Expression per Cluster")) +
    xlab("Cluster Name") +
    ylab("Motif")

  return(new_plot)  
}
```

### Prepare Seurat Object

```{r} 
# Load the object and rename groups
object <- readRDS(object.name)
object$cell_id <- object@active.ident
group.by <- "cell_id"
obj.groups <- GetGroups(
  object = object,
  group.by = group.by,
  idents = NULL
)

# Remove non-EC clusters 
obj.groups.levels <- levels(obj.groups)
obj.groups.levels.EC <- obj.groups.levels[grepl(cell.type, obj.groups.levels)]
object <- subset(object, idents = obj.groups.levels.EC)
```

### Produce Heatmaps

```{r}
plotA <- ProduceHeatmap(
  object = object,
  numberOfMotifs = 15,
  cell.id = obj.groups.levels.EC,
  fold.enrichment = 2.8
)

plotB <- ProduceHeatmap2(
  object = object,
  numberOfMotifs = 25,
  cell.id = obj.groups.levels.EC,
  fold.enrichment = 2.8
)
```
